---
title: "Data Analysis - MSSP Group C2 Fall 2025"
author: "Madeleine Livaudais"
date: "2025-09-23"
output: html_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
set.seed(42)
``` 

## Purpose and Dataset

We will look:

- How one should approach a dataset
- What are the steps followed in analyzing a dataset

Goal. Explore the AirQuality dataset (hourly averages, March 2004–Feb 2005) to:

- understand structure & quality,
- handle missing values,
- do basic EDA (trends, correlations, drift),
- build a simple model for a target variable (e.g., CO).


#### Why these steps?

Good analysis = (1) clean data → (2) understand patterns → (3) model carefully. Each step documents decisions so results are reproducible and defensible.

### Libraries and Project Setup

Use this space to install and store all the packages.

```{r packages, echo=TRUE, message=FALSE}
# Install once if needed and comment it out
# install.packages(c("tidyverse","dplyr", GGally", "ggplot2"))
library(tidyverse)
library(dplyr)
library(GGally) 
library(ggplot2)
```

### Import Data

Read the DataDescription.txt and note important information. Read it thoroughly!

```{r}
aq_raw <- read.csv("AirQualityData.csv", sep = ";")
aq_data <- aq_raw

#Look at the data
head(aq_data)
str(aq_data)
summary(aq_data)

#Replace missing values with NA's
aq_data[aq_data == -200] <- NA

# Rename a single column by its position (e.g., the second column)
names(aq_data)[3] <- "GroundCO"
names(aq_data)[4] <- "SensorCO"

hist(aq_data$GroundCO)

```

